\thetitle

### 2013-11-06 tfpie/eptcs/tfpie.tex
Everything for final EPTCS version.
Copy over eptcs edits.

In the first semester, students are introduced to the ``design
recipe'', a step-by-step process for going from English descriptions
of problems to working programs.  The design recipe involves six steps:
\begin{enumerate}
\item Analyze the information involved in the problem and express
  how to represent it as data.
\item Write down a function signature, a summary of the purporse
  of the function, and a function stub.
\item Illustrate the signature and the purpose statement with some
  functional examples.
\item Take an inventory of the input data that can be used to compute
  an answer.
\item Write the code for the function.
\item Verify the behavior of the program against the functional
  examples given earlier.
\end{enumerate}

Students explore program design using this process in the context of a
series of successively richer pedagogical programming language
levels~\cite{dvanhorn:Felleisen2004Structure,
  dvanhorn:Felleisen2001How} that are included in the DrRacket
(formerly DrScheme) programming
environment~\cite{dvanhorn:Findler2002DrScheme}.
%
These languages are called the beginning student language (BSL), the
intermediate student language (ISL), and the advance student language
(ASL).
%
The language and environment include several tools in support of the
design recipe.  For example, functional examples can be written as
executable tests by writting \texttt{check-expect}
expression~\cite{local:check-expect}; an algebraic stepper and REPL
are provided to interact with programs at each step of the design
process.

@@

\profj also takes on the the dual challenges of implementing Java as
well as subsetting it.  This ultimately resulted in both a limited
Java environment as well as the eventual abandoning of the tool since
it was too difficult to maintain, let alone keep up with advances in
Java.

@@

We have now completed the third iteration of this course, teaching
approximately 35 students each time.  Our experience has been
uniformly positive, and the students have gone on to significant
success in the subsequent courses, despite the curriculum differing
from what the bulk of Northeastern University computer science majors
take.  Anectodally, the class has also had notable success in the
recruitment and retention of female students, as compared to the other
versions of the second-semester course. However, the classes are
sufficiently different as to make a precise comparison impossible.

@@

% \nocite{*}
\bibliographystyle{plainnat}
\bibliography{dvh-bibliography,local,sth-bibliography}

\newpage
\appendix
\section{Worlds and the State pattern}
\label{sec:appendix}

>>>>

In the first semester, students are introduced to the ``design
recipe'', a step-by-step process for going from English descriptions
of problems to working programs.  The design recipe involves six steps:
\begin{enumerate}
\item Analyze the information involved in the problem and express
  how to represent it as data.
\item Write down a function signature, a summary of the purpose
  of the function, and a function stub.
\item Illustrate the signature and the purpose statement with some
  functional examples.
\item Take an inventory of the input data that can be used to compute
  an answer.
\item Write the code for the function.
\item Verify the behavior of the program against the functional
  examples given earlier.
\end{enumerate}

Students explore program design using this process in the context of a
series of successively richer pedagogical programming language
levels~\cite{dvanhorn:Felleisen2004Structure,
  dvanhorn:Felleisen2001How} that are included in the DrRacket
(formerly DrScheme) programming
environment~\cite{dvanhorn:Findler2002DrScheme}.
%
These languages are called the beginning student language (BSL), the
intermediate student language (ISL), and the advance student language
(ASL).
%
The language and environment include several tools in support of the
design recipe.  For example, functional examples can be written as
executable tests by writing \texttt{check-expect}
expression~\cite{local:check-expect}; an algebraic stepper and REPL
are provided to interact with programs at each step of the design
process.

@@

\profj also takes on the dual challenges of implementing Java as
well as subsetting it.  This ultimately resulted in both a limited
Java environment as well as the eventual abandoning of the tool since
it was too difficult to maintain, let alone keep up with advances in
Java.

@@

We have now completed the third iteration of this course, teaching
approximately 35 students each time.  Our experience has been
uniformly positive, and the students have gone on to significant
success in the subsequent courses, despite the curriculum differing
from what the bulk of Northeastern University computer science majors
take.  Anecdotally, the class has also had notable success in the
recruitment and retention of female students, as compared to the other
versions of the second-semester course. However, the classes are
sufficiently different as to make a precise comparison impossible.

@@

% \nocite{*}
\bibliographystyle{eptcs}
\bibliography{dvh-bibliography,local,sth-bibliography}

\appendix
\section{Worlds and the State pattern}
\label{sec:appendix}
